### 2021-09-08 manuscript/BTTB_matrix_vector.tex

This appendix follows the same approach presented by \citet{takahashi2020convolutional}
to efficiently compute the product of a BTTB matrix
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equations \ref{eq:A_expand} and \ref{eq:BTTB_A_alpha_beta}) 
and a generic vector $\mathbf{b}$. Let this matrix-vector product be represented by
\begin{equation}
\mathbf{t}_{\boldsymbol{\alpha\beta}} = 
\mathbf{A_{\boldsymbol{\alpha\beta}}} \, \mathbf{b} \: ,
\label{eq:t-alpha-beta}
\end{equation}
where
\begin{equation}
\mathbf{t}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{t}^{0}_{\boldsymbol{\alpha\beta}} \\
\vdots \\
\mathbf{t}^{Q-1}_{\boldsymbol{\alpha\beta}} \\
\end{bmatrix}_{N \times 1} \: ,
\label{eq:t-alpha-beta-partitioned}
\end{equation} 
\begin{equation}
\mathbf{b} = \begin{bmatrix}
\mathbf{b}^{0} \\
\vdots \\
\mathbf{b}^{Q-1} \\
\end{bmatrix}_{N \times 1} \: ,
\label{eq:b-partitioned}
\end{equation}
$\mathbf{t}^{q}_{\boldsymbol{\alpha\beta}}$ and $\mathbf{b}^{q}$ are $P \times 1$ vectors 
and $q$ is the block index (equations \ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}). 
From equation \ref{eq:t-alpha-beta}, we obtain an auxiliary matrix-vector product given by
\begin{equation}
\mathbf{w}_{\boldsymbol{\alpha\beta}} = \mathbf{C}_{\boldsymbol{\alpha\beta}} \, \mathbf{v} \: ,
\label{eq:w_alpha_beta}
\end{equation}
where
\begin{equation}
\mathbf{w}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{w}_{\boldsymbol{\alpha\beta}}^{0} \\
\vdots \\
\mathbf{w}_{\boldsymbol{\alpha\beta}}^{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \quad ,
\label{eq:w_alpha_beta_partitioned}
\end{equation}
\begin{equation}
\mathbf{w}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
\mathbf{t}^{q}_{\boldsymbol{\alpha\beta}} \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1}
\label{eq:wq-vector} \quad ,
\end{equation}
\begin{equation}
\mathbf{v} = \begin{bmatrix}
\mathbf{v}^{0} \\
\vdots \\
\mathbf{v}^{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \: ,
\label{eq:v-vector}
\end{equation}
\begin{equation}
\mathbf{v}^{q} = \begin{bmatrix}
\mathbf{b}^{q} \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1} \: ,
\label{eq:vq-vector} 
\end{equation}
$\mathbf{0}_{2N \times 1}$ and $\mathbf{0}_{P \times 1}$ are vectors of zeros
and $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is a $4N \times 4N$ 
block circulant matrix with circulant blocks (BCCB) \citep[e.g., ][ p. 184]{davis1979}.
The key point here is that the auxiliary matrix-vector product
(equation \ref{eq:w_alpha_beta}) represents a 2D discrete convolution and can be 
efficiently computed by using the 2D Fast Fourier Transform (2D FFT).

The BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is formed by $2Q \times 2Q$ blocks, 
where each block $\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$ is a $2P \times 2P$ circulant matrix.
The entire BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is defined by properly
downshifting its first block column
\begin{equation}
	\left[ \mathbf{C}_{\boldsymbol{\alpha\beta}} \right]_{(0)} = \begin{bmatrix}
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{0} \\
		\vdots \\
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{Q-1} \\
		\mathbf{0}_{2P \times 2P} \\
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{-Q+1} \\
		\vdots \\
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{-1}
	\end{bmatrix}_{4N \times 2P} \: ,
	\label{eq:C_alpha_beta_first_block_column}
\end{equation}
where $\mathbf{0}_{2P \times 2P}$ is a matrix of zeros and 
each block $\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$, $q = -Q+1, \dots, Q-1$,
is obtained by properly downshifting its first column
\begin{equation}
	\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
		a^{\alpha\beta}_{q0} \\
		\vdots \\
		a^{\alpha\beta}_{q(P-1)} \\
		0 \\
		a^{\alpha\beta}_{q(-P+1)} \\
		\vdots \\
		a^{\alpha\beta}_{q(-1)}
	\end{bmatrix}_{2P \times 1} \: ,
	\label{eq:cq_alpha_beta}
\end{equation}
where $a^{\alpha\beta}_{qp}$, $p = -P+1, \dots, P-1$, are the elements of
matrix component $\mathbf{A_{\boldsymbol{\alpha\beta}}}$. 
The BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is diagonalized by
$\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P}$, where ``$\otimes$" denotes the Kronecker product
\citep[e.g.,][ p. 242]{horn_johnson1991} and $\mathbf{F}_{2Q}$ and $\mathbf{F}_{2P}$ are 
the $2Q \times 2Q$ and $2P \times 2P$ unitary DFT matrices \citep[][ p. 31]{davis1979}.
Due to this property, the auxiliary matrix-vector product (equation \ref{eq:w_alpha_beta}) 
can be computed as follows \citep{takahashi2020convolutional}:
\begin{equation}
\mathbf{F}_{2Q}^{\ast} \left[ 
\mathbf{L}_{\boldsymbol{\alpha\beta}} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
\right] \mathbf{F}_{2P}^{\ast} = \mathbf{W}_{\boldsymbol{\alpha\beta}} \: ,
\label{eq:2d-discrete-convolution}
\end{equation}
where ``$\circ$" denotes the Hadamard (element-wise) product \citep[e.g.,][ p. 298]{horn_johnson1991},
``$\ast$" denotes the complex conjugate, 
$\mathbf{W}_{\boldsymbol{\alpha\beta}}$ and $\mathbf{V}$ are $2Q \times 2P$ matrices obtained
by rearranging, respectively, vectors $\mathbf{w}_{\boldsymbol{\alpha\beta}}$ 
(equation \ref{eq:w_alpha_beta_partitioned}) and $\mathbf{v}$ (equation \ref{eq:v-vector})
along their rows and
\begin{equation}
\mathbf{L}_{\boldsymbol{\alpha\beta}} = \sqrt{4QP} \; 
\mathbf{F}_{2Q} \, \mathbf{G}_{\boldsymbol{\alpha\beta}} \, \mathbf{F}_{2P} \: ,
\label{eq:L_alpha_beta}
\end{equation}
where
\begin{equation}
\mathbf{G}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\left( \mathbf{c}_{\boldsymbol{\alpha\beta}}^{0} \right)^{\top} \\
\vdots \\
\left( \mathbf{c}_{\boldsymbol{\alpha\beta}}^{Q-1} \right)^{\top} \\
\mathbf{0}_{Q \times 2P}
\end{bmatrix}_{2Q \times 2P} \: ,
\label{eq:G_alpha_beta}
\end{equation}
with $\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q}$, $q = -Q+1, \dots, Q-1$, given by equation
\ref{eq:cq_alpha_beta}.
Note that matrix $\mathbf{G}_{\boldsymbol{\alpha\beta}}$ 
(equation \ref{eq:G_alpha_beta}) is formed by the first column 
$\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q}$ (equation \ref{eq:cq_alpha_beta}) of each 
circulant block $\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$. It means that all elements
of matrix $\mathbf{G}_{\boldsymbol{\alpha\beta}}$ can be obtained by using only the first 
column of $\mathbf{A_{\boldsymbol{\alpha\beta}}}$.

It is important noting that the left side of equation \ref{eq:2d-discrete-convolution} represents 
the 2D Inverse Discrete Fourier Transform (2D IDFT) of the term in brackets. This term, in turn,
represents the Hadamard product of $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ 
(equation \ref{eq:L_alpha_beta}) and the 2D Discrete Fourier Transform (2D DFT) of $\mathbf{V}$.
Similarly, equation \ref{eq:L_alpha_beta} shows that $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ is
obtained by computing the 2D DFT of matrix $\mathbf{G}_{\boldsymbol{\alpha\beta}}$
(equation \ref{eq:G_alpha_beta}). Hence, equations \ref{eq:2d-discrete-convolution} and
\ref{eq:L_alpha_beta} can be efficiently computed by using the 2D FFT. After that, the elements of
vector $\mathbf{t}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:t-alpha-beta}) can be
retrieved from the first quadrant of matrix $\mathbf{W}_{\boldsymbol{\alpha\beta}}$
(equation \ref{eq:2d-discrete-convolution}).

Now, consider the matrix-vector product
\begin{equation}
\mathbf{t} = 
\mathbf{A} \, \mathbf{b} \: ,
\label{eq:t}
\end{equation}
where 
\begin{equation}
\mathbf{t} = \mathbf{t}_{\boldsymbol{xx}} + \mathbf{t}_{\boldsymbol{xy}} + \mathbf{t}_{\boldsymbol{xz}} +
\mathbf{t}_{\boldsymbol{yy}} + \mathbf{t}_{\boldsymbol{yz}} + \mathbf{t}_{\boldsymbol{zz}}
\label{eq:t-components}
\end{equation}
and $\mathbf{A}$ is the sensitivity matrix given by equation \ref{eq:A_expand}. It can be shown
that this product can be efficiently computed by
\begin{equation}
\mathbf{F}_{2Q}^{\ast} \left[ 
\mathbf{L} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
\right] \mathbf{F}_{2P}^{\ast} = \mathbf{W} \: ,
\label{eq:2d-discrete-convolution-complete}
\end{equation}
where
\begin{equation}
\mathbf{W} = \mathbf{W}_{\boldsymbol{xx}} + \mathbf{W}_{\boldsymbol{xy}} + \mathbf{W}_{\boldsymbol{xz}} + \mathbf{W}_{\boldsymbol{yy}} + \mathbf{W}_{\boldsymbol{yz}} + \mathbf{W}_{\boldsymbol{zz}} \: ,
\label{eq:W}
\end{equation}
\begin{equation}
\mathbf{L} = \mathbf{L}_{\boldsymbol{xx}} + \mathbf{L}_{\boldsymbol{xy}} + \mathbf{L}_{\boldsymbol{xz}} + \mathbf{L}_{\boldsymbol{yy}} + \mathbf{L}_{\boldsymbol{yz}} + \mathbf{L}_{\boldsymbol{zz}}
\label{eq:L}
\end{equation}
and $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ is given by equation \ref{eq:L_alpha_beta}.
Finally, it can be shown that these equation can also be used to compute the 
matrix-vector product of $\mathbf{A}$ (equation \ref{eq:A_expand}) and an arbitrary vector 
$\mathbf{b}$. The only difference is that, in this case, each circulant block 
$\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$, $q = -Q+1, \dots, Q-1$, is obtained by properly
downshifting the modified vector
\begin{equation}
\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
a^{\alpha\beta}_{q0} \\
\vdots \\
a^{\alpha\beta}_{q(-P+1)} \\
0 \\
a^{\alpha\beta}_{q(P-1)} \\
\vdots \\
a^{\alpha\beta}_{q1}
\end{bmatrix}_{2P \times 1} \: .
\label{eq:cq_alpha_beta_trnasposed}
\end{equation}

>>>>

This appendix follows a similar approach to that presented by \citet{takahashi2020convolutional}
to efficiently compute the product of the sensitivity matrix
$\mathbf{A}$ (equations \ref{eq:A_expand}) and a generic vector $\mathbf{b}$. 
Let this product be represented by
\begin{equation}
	\mathbf{t} = 
	\mathbf{A} \, \mathbf{b} \: ,
	\label{eq:t}
\end{equation}
where 
\begin{equation}
	\mathbf{t} = \mathbf{t}_{\boldsymbol{xx}} + \mathbf{t}_{\boldsymbol{xy}} + \mathbf{t}_{\boldsymbol{xz}} +
	\mathbf{t}_{\boldsymbol{yy}} + \mathbf{t}_{\boldsymbol{yz}} + \mathbf{t}_{\boldsymbol{zz}}
	\label{eq:t-components}
\end{equation}
and
\begin{equation}
\mathbf{t}_{\boldsymbol{\alpha\beta}} = 
\mathbf{A_{\boldsymbol{\alpha\beta}}} \, \mathbf{b} \: .
\label{eq:t-alpha-beta}
\end{equation}
Let us also consider that vectors
\begin{equation}
\mathbf{t}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{t}^{0}_{\boldsymbol{\alpha\beta}} \\
\vdots \\
\mathbf{t}^{Q-1}_{\boldsymbol{\alpha\beta}} \\
\end{bmatrix}_{N \times 1}
\label{eq:t-alpha-beta-partitioned}
\end{equation} 
and
\begin{equation}
\mathbf{b} = \begin{bmatrix}
\mathbf{b}^{0} \\
\vdots \\
\mathbf{b}^{Q-1} \\
\end{bmatrix}_{N \times 1}
\label{eq:b-partitioned}
\end{equation}
are composed of $P \times 1$ vectors $\mathbf{t}^{q}_{\boldsymbol{\alpha\beta}}$ and $\mathbf{b}^{q}$,
respectively, where $q$ is the block index (equations \ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}). 
From equation \ref{eq:t-alpha-beta}, we obtain an auxiliary matrix-vector product given by
\begin{equation}
\mathbf{w}_{\boldsymbol{\alpha\beta}} = \mathbf{C}_{\boldsymbol{\alpha\beta}} \, \mathbf{v} \: ,
\label{eq:w_alpha_beta}
\end{equation}
where $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is a $4N \times 4N$ 
block circulant matrix with circulant blocks (BCCB) \citep[e.g., ][ p. 184]{davis1979},
\begin{equation}
\mathbf{w}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{w}_{\boldsymbol{\alpha\beta}}^{0} \\
\vdots \\
\mathbf{w}_{\boldsymbol{\alpha\beta}}^{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \quad ,
\label{eq:w_alpha_beta_partitioned}
\end{equation}
\begin{equation}
\mathbf{w}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
\mathbf{t}^{q}_{\boldsymbol{\alpha\beta}} \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1}
\label{eq:wq-vector} \quad ,
\end{equation}
\begin{equation}
\mathbf{v} = \begin{bmatrix}
\mathbf{v}^{0} \\
\vdots \\
\mathbf{v}^{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1}
\label{eq:v-vector}
\end{equation}
and
\begin{equation}
\mathbf{v}^{q} = \begin{bmatrix}
\mathbf{b}^{q} \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1} \: ,
\label{eq:vq-vector} 
\end{equation}
with $\mathbf{0}_{2N \times 1}$ and $\mathbf{0}_{P \times 1}$ being vectors of zeros.
The key point here is that the auxiliary matrix-vector product
(equation \ref{eq:w_alpha_beta}) represents a 2D discrete convolution and can be 
efficiently computed by using the 2D Fast Fourier Transform (2D FFT).

The BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta})
is formed by $2Q \times 2Q$ blocks, 
where each block $\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$ is a $2P \times 2P$ circulant matrix.
The entire BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is defined by properly
downshifting its first block column
\begin{equation}
	\left[ \mathbf{C}_{\boldsymbol{\alpha\beta}} \right]_{(0)} = \begin{bmatrix}
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{0} \\
		\vdots \\
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{Q-1} \\
		\mathbf{0}_{2P \times 2P} \\
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{-Q+1} \\
		\vdots \\
		\mathbf{C}_{\boldsymbol{\alpha\beta}}^{-1}
	\end{bmatrix}_{4N \times 2P} \: ,
	\label{eq:C_alpha_beta_first_block_column}
\end{equation}
where $\mathbf{0}_{2P \times 2P}$ is a matrix of zeros. Similarly, 
each block $\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$, $q = -Q+1, \dots, Q-1$,
is obtained by properly downshifting its first column
\begin{equation}
	\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
		a^{\alpha\beta}_{q0} \\
		\vdots \\
		a^{\alpha\beta}_{q(P-1)} \\
		0 \\
		a^{\alpha\beta}_{q(-P+1)} \\
		\vdots \\
		a^{\alpha\beta}_{q(-1)}
	\end{bmatrix}_{2P \times 1} \: ,
	\label{eq:cq_alpha_beta}
\end{equation}
where $a^{\alpha\beta}_{qp}$, $p = -P+1, \dots, P-1$, are the elements of
matrix component $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ described in terms of
block indices $q$ and $p$ (equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}). 
The BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ is diagonalized by
$\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P}$, where ``$\otimes$" denotes the Kronecker product
\citep[e.g.,][ p. 242]{horn_johnson1991} and $\mathbf{F}_{2Q}$ and $\mathbf{F}_{2P}$ are 
the $2Q \times 2Q$ and $2P \times 2P$ unitary DFT matrices \citep[][ p. 31]{davis1979}.
Due to this property, the auxiliary matrix-vector product (equation \ref{eq:w_alpha_beta}) 
can be computed as follows \citep{takahashi2020convolutional}:
\begin{equation}
\mathbf{F}_{2Q}^{\ast} \left[ 
\mathbf{L}_{\boldsymbol{\alpha\beta}} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
\right] \mathbf{F}_{2P}^{\ast} = \mathbf{W}_{\boldsymbol{\alpha\beta}} \: ,
\label{eq:2d-discrete-convolution}
\end{equation}
where ``$\circ$" denotes the Hadamard (element-wise) product \citep[e.g.,][ p. 298]{horn_johnson1991},
``$\ast$" denotes the complex conjugate, 
$\mathbf{W}_{\boldsymbol{\alpha\beta}}$ and $\mathbf{V}$ are $2Q \times 2P$ matrices obtained
by rearranging, respectively, vectors $\mathbf{w}_{\boldsymbol{\alpha\beta}}$ 
(equation \ref{eq:w_alpha_beta_partitioned}) and $\mathbf{v}$ (equation \ref{eq:v-vector})
along their rows and $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ is a $2Q \times 2P$ matrix
given by
\begin{equation}
\mathbf{L}_{\boldsymbol{\alpha\beta}} = \sqrt{4QP} \; 
\mathbf{F}_{2Q} \, \mathbf{G}_{\boldsymbol{\alpha\beta}} \, \mathbf{F}_{2P} \: ,
\label{eq:L_alpha_beta}
\end{equation}
with
\begin{equation}
\mathbf{G}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\left( \mathbf{c}_{\boldsymbol{\alpha\beta}}^{0} \right)^{\top} \\
\vdots \\
\left( \mathbf{c}_{\boldsymbol{\alpha\beta}}^{Q-1} \right)^{\top} \\
\mathbf{0}_{1 \times 2P} \\
\left( \mathbf{c}_{\boldsymbol{\alpha\beta}}^{-Q+1} \right)^{\top} \\
\vdots \\
\left( \mathbf{c}_{\boldsymbol{\alpha\beta}}^{-1} \right)^{\top}
\end{bmatrix}_{2Q \times 2P} \: ,
\label{eq:G_alpha_beta}
\end{equation}
defined by the first columns $\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q}$ 
(equation \ref{eq:cq_alpha_beta}), $q = -Q+1, \dots, Q-1$, of all circulant blocks
$\mathbf{C}_{\boldsymbol{\alpha\beta}}^{q}$ (equation \ref{eq:C_alpha_beta_first_block_column}).
Hence, the whole BCCB matrix $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ does not have to be
formed, but only its first column. 
Besides, the symmetries defined by equations 
\ref{eq:Axx_symmetry}--\ref{eq:Ayz_q_internal_block_symmetry_y_oriented} imply
that all elements of $\mathbf{G}_{\boldsymbol{\alpha\beta}}$ can be obtained by 
using only the first column of $\mathbf{A_{\boldsymbol{\alpha\beta}}}$. Consequently, 
the whole matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ do not have to be formed
as well, but only their first columns.

It is important noting that the left side of equation \ref{eq:2d-discrete-convolution} represents 
the 2D Inverse Discrete Fourier Transform (2D IDFT) of the term in brackets. This term, in turn,
represents the Hadamard product of $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ 
(equation \ref{eq:L_alpha_beta}) and the 2D Discrete Fourier Transform (2D DFT) of $\mathbf{V}$.
Similarly, equation \ref{eq:L_alpha_beta} shows that $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ is
obtained by computing the 2D DFT of matrix $\mathbf{G}_{\boldsymbol{\alpha\beta}}$
(equation \ref{eq:G_alpha_beta}). Hence, equations \ref{eq:2d-discrete-convolution} and
\ref{eq:L_alpha_beta} can be efficiently computed by using the 2D FFT. After that, the elements of
vector $\mathbf{t}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:t-alpha-beta}) can be
retrieved from the first quadrant of matrix $\mathbf{W}_{\boldsymbol{\alpha\beta}}$
(equation \ref{eq:2d-discrete-convolution}). By combining the results obtained for all
components $\boldsymbol{\alpha\beta}$, 
$\boldsymbol{\alpha}, \boldsymbol{\beta} = \boldsymbol{x}, \boldsymbol{y}, \boldsymbol{z}$,
we can show that
\begin{equation}
\mathbf{F}_{2Q}^{\ast} \left[ 
\mathbf{L} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
\right] \mathbf{F}_{2P}^{\ast} = \mathbf{W} \: ,
\label{eq:2d-discrete-convolution-complete}
\end{equation}
where
\begin{equation}
\mathbf{W} = \mathbf{W}_{\boldsymbol{xx}} + \mathbf{W}_{\boldsymbol{xy}} + \mathbf{W}_{\boldsymbol{xz}} + \mathbf{W}_{\boldsymbol{yy}} + \mathbf{W}_{\boldsymbol{yz}} + \mathbf{W}_{\boldsymbol{zz}}
\label{eq:W}
\end{equation}
and
\begin{equation}
\mathbf{L} = \mathbf{L}_{\boldsymbol{xx}} + \mathbf{L}_{\boldsymbol{xy}} + \mathbf{L}_{\boldsymbol{xz}} + \mathbf{L}_{\boldsymbol{yy}} + \mathbf{L}_{\boldsymbol{yz}} + \mathbf{L}_{\boldsymbol{zz}} \: ,
\label{eq:L}
\end{equation}
with $\mathbf{L}_{\boldsymbol{\alpha\beta}}$ defined by equation \ref{eq:L_alpha_beta}.
Then, the elements of $\mathbf{t}$ (equation \ref{eq:t}) are obtained from the first
quadrant of $\mathbf{W}$ (equations \ref{eq:2d-discrete-convolution-complete} and
\ref{eq:W}).

Finally, it can be shown that the product 
\begin{equation}
\mathbf{t} = 
\mathbf{A}^{\top} \mathbf{b}
\label{eq:t_AT}
\end{equation}
can be computed by using equation \ref{eq:2d-discrete-convolution-complete}.
The difference is that, in this case, matrices $\mathbf{G}_{\boldsymbol{\alpha\beta}}$
(equation \ref{eq:G_alpha_beta}) are defined by using the new vectors
\begin{equation}
\mathbf{c}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
a^{\alpha\beta}_{q0} \\
\vdots \\
a^{\alpha\beta}_{q(-P+1)} \\
0 \\
a^{\alpha\beta}_{q(P-1)} \\
\vdots \\
a^{\alpha\beta}_{q1}
\end{bmatrix}_{2P \times 1} \: .
\label{eq:cq_alpha_beta_trnasposed}
\end{equation}

### 2021-09-08 manuscript/abstract.tex

A fast equivalent layer for magnetic data processing is presented. Taking advantage of the sensitivity matrix structure of the magnetic kernel, when observation and equivalent sources are aligned on a regular spaced grid with constant height, it is possible to calculate the matrix-vector product in a very fast manner. The structure is called block-Toeplitz Toeplitz-block (BTTB) and this type of matrix is well known in literature to be embbeded in a block-Circulant Circulant-block (BCCB), wich in turn can have its eigenvalues calculated using only the first column and a 2D fast Fourier transform. We show that, despite this BTTB matrix is not symmetric, by using only the first equivalent source it is possible to calculate all the first columns of the inverse of distance second derivatives that composes the magnetic kernel and reconstruct the first column of the BCCB matrix, saving computational time and system memory. The conjugate gradient iterative method is used to solve the linear system and estimate the physical properties distributed over the equivalent layer. Synthetic tests show a decrease in the order of $10^4$ in floating-point operations, $25$ times in computation runtime with a mid-size $80 \times 80$ grid, and exponential decrease in memory RAM usage, allowing to perform this operation with millions of observation points on desktop computers. Synthetic tests with irregular grids also show that this method can work with directional disturbances under certain limits. A real magnetic data of Caraj{\'a}s Province, Brazil, with $1,310,000$ observation points in an irregular grid was used to successfully perform a data processing with this method, taking $385.56$ seconds to estimate the physical property and $2.64$ seconds for the upward-continuation.

>>>>

We present a fast equivalent layer for processing large-scale magnetic data. 
Taking advantage of the sensitivity matrix structure of the magnetic kernel, when observation and equivalent sources are aligned on a regular spaced grid with constant height, we can calculate the matrix-vector product in a very fast manner. 
In the literature, this matrix structure is called block-Toeplitz Toeplitz-block (BTTB) that can be embbeded 
into a block-Circulant Circulant-block (BCCB), which in turn has its eigenvalues calculated using only the first column of the BTTB matrix and a 2D fast Fourier transform. 
We demonstrate that, despite this BTTB matrix is not symmetric, by using only the first equivalent source it is possible to calculate all the first columns of the inverse of distance second derivatives that compose the magnetic kernel and reconstruct the first column of the BCCB matrix, saving computational time and system memory. 
We solve the linear system by modifying the conjugate gradient iterative method to estimate the 
physical-property distribution over the equivalent layer and then to process large data sets.
Synthetic tests show a decrease in the order of $10^4$ in floating-point operations, $25$ times in computation runtime with a mid-size $80 \times 80$ grid, and exponential decrease in memory RAM usage, allowing to perform this operation with millions of observations on desktop computers. 
Synthetic tests simulating data on irregular grids or over undulating observation surfaces  show the robustness of the convolutional equivalent layer in processing magnetic surveys that violate the requirement  that the data be measured on a regular grid and the observation surface be planar.
Test on real magnetic data from Caraj{\'a}s Province, Brazil, with $1,310,000$ observations on an irregular grid, confirms the success of our method, taking $385.56$ seconds to estimate the physical-property distribution over the equivalent layer and $2.64$ seconds for upward-continuing the data.

### 2021-09-08 manuscript/tables.tex
Tabelas 2,3 e 4 com as medias e desvios padroes dos residuos testes sinteticos e reais
Medias e desvios padroes dos residuos atualizaods com a unidade de medida
Novos resultados dos dados reais com a inclinacao calculada pela calculadora de campo magnetico da NOAA.

\tabl{RAM-usage}{This table shows the RAM memory usage (in Megabytes) for storing the whole matrix $\mathbf{A}$ (equation \ref{eq:aij_mag}), the sum of all six first columns of the BCCB matrices embedded from the components of the matrix $\mathbf{H}$ from equation \ref{eq:Hi} (both need 8 bytes per element) and the matrix $\mathbf{L}$ containing the eigenvalues complex numbers (16 bytes per element) resulting from the diagonalization of matrix $\mathbf{C}$ (equation \ref{eq:w_Cv}). Here we must consider that $N$ observation points forms a $N \times N$ matrix.
\label{tab:RAM-usage}}
{
	\begin{center}
		\begin{tabular}[]{|l|c|c|c|}
			\hline
			\textbf{$N$} & \textbf{Matrix $\mathbf{A}$} & \textbf{All six first columns of BCCB matrices} & \textbf{Matrix $\mathbf{L}$}\\
			\hline 
			$100$ & 0.0763 & 0.0183 & 0.00610\\
			\hline
			$400$ & 1.22 & 0.0744 & 0.0248\\
			\hline
			$2,500$ & 48 & 0.458 & 0.1528\\
			\hline
			$10,000$ & 763 & 1.831 & 0.6104\\
			\hline
			$40,000$ & 12,207 & 7.32 & 2.4416 \\
			\hline
			$250,000$ & 476,837 & 45.768 & 15.3 \\
			\hline
			$500,000$ & 1,907,349 & 91.56 & 30.518 \\
			\hline
			$1,000,000$ & 7,629,395 & 183.096 & 61.035 \\
			\hline
		\end{tabular}
	\end{center} 
}

>>>>

\tabl{RAM-usage}{This table shows the RAM memory usage (in Megabytes) for storing the whole 
	$N \times N$ matrix $\mathbf{A}$ (equation \ref{eq:A_expand}), the first columns of the BCCB
	matrices $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) (both need 8 
	bytes per element) and the matrix $\mathbf{L}$ (equation \ref{eq:L}) (16 bytes per element).
\label{tab:RAM-usage}}
{
	\begin{center}
		\begin{tabular}[]{|l|c|c|c|}
			\hline
			\textbf{$N$} & \textbf{$\mathbf{A}$} & \textbf{First columns of matrices $\mathbf{C}_{\boldsymbol{\alpha\beta}}$} & \textbf{$\mathbf{L}$}\\
			\hline 
			$100$ & 0.0763 & 0.0183 & 0.00610\\
			\hline
			$400$ & 1.22 & 0.0744 & 0.0248\\
			\hline
			$2,500$ & 48 & 0.458 & 0.1528\\
			\hline
			$10,000$ & 763 & 1.831 & 0.6104\\
			\hline
			$40,000$ & 12,207 & 7.32 & 2.4416 \\
			\hline
			$250,000$ & 476,837 & 45.768 & 15.3 \\
			\hline
			$500,000$ & 1,907,349 & 91.56 & 30.518 \\
			\hline
			$1,000,000$ & 7,629,395 & 183.096 & 61.035 \\
			\hline
		\end{tabular}
	\end{center} 
}
